\documentclass{article}

\usepackage{amsmath, amssymb, amsthm}

\usepackage{logicpuzzle}

% YYYY-MM-DD Date format
\usepackage[style=iso]{datetime2}

% Allows use of internal commands
\makeatletter

% Place an X for an empty cell
\newcommand*\placex[2]{
    \LP@ingrid{#1}{#2}{\LP@NG@columns}{\LP@NG@rows}{nonogram}
    \LP@G@setcellcontent[hcenter,vcenter]{#1}{#2}{$\times$}
    }

% For filling a solved puzzle's empty squares with X marks
% Current solution is to fill every cell with an X and filled cells will cover it.
% Partially finished puzzles will need each desired X placed individually.
% TODO is there anything less stupid
\newcommand{\nonogramfill}[2]{
    \foreach \i in {1,...,#1}
    \foreach \j in {1,...,#2}
    {
        \placex{\i}{\j}
    }
    }


\title{Nonogram Worksheet}
\author{Michael Gadhia}
\date{\today}

\begin{document}
\maketitle

% What this document is and what it assumes
\section{Introduction}

This is an annotated worksheet of nonogram/picross puzzles, and assumes basic knowledge of how to solve one.
This will hopefully become a companion to the nonogram fact sheet once it's finished, but I have a lot of TikZ learning to do before I can make nonogram variants with the \texttt{logicpuzzle} package.
This sheet will start with a few simple nonograms and then include ones that aren't line-solvable, and hopefully one day will have color and isometric triangular puzzles (if I can typeset them in \LaTeX).

\section{Normal Puzzles}

Puzzles in this section have a single unique solution you can reach without guessing, and follow the regular rules of nonograms.

This first one was the example in my fact sheet.

\begin{center}
    % 5x5 dog
    \begin{nonogram}[rows=5,columns=5,scale=0.35,width=3cm,
        fontsize=footnotesize,helplines=5,extracells=2]

        \nonogramV{{1,1},{4},{1,2},{1},{1}}
        \nonogramH{{1},{4},{1,1},{1},{3}}

    \end{nonogram}
\end{center}

Here are a few more like that, hopefully solving them lets you develop some techniques.

\begin{center}
    % 5x5 smile
    \begin{nonogram}[rows=5,columns=5,scale=0.35,width=3cm,
        fontsize=footnotesize,helplines=5,extracells=2]

        \nonogramV{{5},{1,1},{1},{2,2},{2,2}}
        \nonogramH{{2,2},{1,2},{1,1},{1,2},{2,2}}

    \end{nonogram}
    \hspace{1cm}
    % 5x5 tree
    \begin{nonogram}[rows=5,columns=5,scale=0.35,width=3cm,
        fontsize=footnotesize,helplines=5,extracells=2]

        \nonogramV{{5},{1},{5},{3},{1}}
        \nonogramH{{1,1},{1,2},{5},{1,2},{1,1}}

    \end{nonogram}

    % 10x10 music note
    \begin{nonogram}[rows=10,columns=10,scale=0.35,width=3cm,
        fontsize=footnotesize,helplines=5,extracells=4]

        \nonogramV{{2},{2,4},{4,4},{4,3},{3,1},{1,1},{1,4},{3,1},{1,3},{4}}
        \nonogramH{{2},{4},{4},{8},{1,1},{1,1},{2,1,1},{4,1,1},{4,1,1},{8}}

    \end{nonogram}
    \hspace{4cm}
    % 10x10 cat
    \begin{nonogram}[rows=10,columns=10,scale=0.35,width=3cm,
        fontsize=footnotesize,helplines=5,extracells=4]

        \nonogramV{{6},{8},{2,6},{2,5},{1,3},{1,1},{1,1,5},{1,1,1,1},{5},{1,1}}
        \nonogramH{{3,4},{5,1,1},{9},{5,1,1},{4,4},{3},{2},{2,1},{2,1},{4}}
        \puzzlestrut

    \end{nonogram}
\end{center}

\subsection{Line-Solvability}

Looking at these next puzzles one line at a time is not enough to solve them.
But, they still have a unique solution, even if they seem confusing at first.
Consider this small example first.

\begin{center}
    % 4x2 curve
    \begin{nonogram}[rows=4,columns=2,scale=0.35,width=2cm,
        fontsize=footnotesize,helplines=5,extracells=2]

        \nonogramV{{1},{1},{1},{1}}
        \nonogramH{{2},{1,1}}

    \end{nonogram}
\end{center}

Did solving that feel different than the previous ones?
Sometimes, you can progress using contradictions.
Imagine one cell being filled, then figure out if assuming that would lead to something impossible.
If it would, that assumption must be wrong, so the cell is empty.
All of these next ones will require multi-line logic at some point, and finding the right assumption to create a contradiction can be slow.
Hopefully my fact sheet will eventually contain advanced techniques that make that easier.

The puzzle on the left is not really a picture of anything, and it may be more difficult.
Note that rows and columns with a 0 have only empty cells, so you can solve those immediately.
If you're stuck after that, think about the top left and try to find something that would be impossible.

The puzzle on the right may be easy to start, but it will still need advanced logic to finish.
If you choose a good cell to assume is or isn't filled, you might only have to find one contradiction.

\begin{center}
    % 5x5 https://puzzling.stackexchange.com/a/129850
    \begin{nonogram}[rows=5,columns=5,scale=0.35,width=3cm,
        fontsize=footnotesize,helplines=5,extracells=2]

        \nonogramV{{1},{2},{0},{2},{2}}
        \nonogramH{{2},{1},{0},{2},{2}}

    \end{nonogram}
    \hspace{1cm}
    % 5x5 https://puzzling.stackexchange.com/a/129931
    \begin{nonogram}[rows=5,columns=5,scale=0.35,width=3cm,
        fontsize=footnotesize,helplines=5,extracells=2]

        \nonogramV{{1,1},{1,2},{2},{2},{1}}
        \nonogramH{{2},{2},{2},{2},{2}}

    \end{nonogram}
\end{center}





\end{document}